---
title: "Final Project Combine Analysis"
author: "Christina Ling, Charles Naylor, Sylvia Mawusi Ayikpah"
date: "December 5, 2018"
output: html_document
---


```{r libraries}
## ONCE: install.packages("tm")
suppressPackageStartupMessages(library(tm))
suppressPackageStartupMessages(library(tidyverse))
library(ggplot2)
library(tokenizers)
library(cluster)
suppressPackageStartupMessages(library(factoextra))
suppressPackageStartupMessages(library(gridExtra))
```

A little bit of cleaning up.

```{r import data and examination}
filename="data.csv"

music <- read.csv(filename, sep=",", header=T, na.string=c(""), stringsAsFactors=F)

str(music)

col.names <- colnames(music)
col.names <- col.names[-1]
col.names <- tolower(col.names)
col.names <- c("rank", col.names)
colnames(music) <- col.names

#check work
str(music)


#add spotify data
filename = "bb_spotify.csv"
spotify <- read.csv(filename, sep=",", header=T, na.string=c(""), stringsAsFactors=F)

head(spotify)
#remove first row
spotify <- spotify[-1,]

str(spotify)
```

The lyrics were transformed into word frequency matrix. The analyze will be looking at how often a word appears in the lyrics. In order to make meaninful analyze, all words were transformed to lower case and all stop-words (commonly used words) were removed. Additionally, all punctuations and numbers were removed. The data frames shows for each songs how frequent a word appears in that song, with the rows being songs and columns being words.

``` {r lyrics word frequency matrix}
lyrics <- c(lyrics = music$lyrics)

corpus <- Corpus(VectorSource(lyrics))
  # remove punctuation
cleancorpus <- tm_map(corpus, removePunctuation)
  # change to lower case
cleancorpus <- tm_map(cleancorpus, content_transformer(tolower))
  # remove numbers
cleancorpus <- tm_map(cleancorpus, removeNumbers)
  # remove whitespace
cleancorpus <- tm_map(cleancorpus, stripWhitespace)
  # remove stop words
stop <- stopwords("english")
stop <- Corpus(VectorSource(stop))
stop <- suppressMessages(tm_map(stop, removePunctuation))
stop <- unlist(stop)
cleancorpus <- tm_map (cleancorpus, removeWords, stop)

tdm <- TermDocumentMatrix(cleancorpus)

lyrics <- as.matrix(tdm)
  #change to data frame
lyrics <- as.data.frame(lyrics)
lyrics <- as.data.frame(t(lyrics))

# look at results
  lyrics[1:7,1:7]

  #change row names
rownames(lyrics) <- c(1:5100)
```

Combining the data into one

``` {R combine data}
index <- grep("featuring", music$artist) +1

featuring <- replicate(5100, 0)
featuring[index] = 1

df <- data.frame(rank = music$rank,
                 danceability = spotify$danceability,
                 energy = spotify$energy,
                 loudness = spotify$loudness,
                 speechiness = spotify$speechiness,
                 acousticness = spotify$acousticness,
                 instrmentalness = spotify$instrumentalness,
                 liveness = spotify$liveness,
                 tempo = spotify$temp,
                 duration = spotify$duration_ms,
                 girl = lyrics$girl,
                 love = lyrics$love, 
                 ooh = lyrics$ooh,
                 want = lyrics$want,
                 baby = lyrics$baby,
                 yeah = lyrics$yeah,
                 back = lyrics$back,
                 cause = lyrics$cause,
                 thing = lyrics$thing,
                 stringsAsFactors = F)

for (column in 1:length(colnames(df))) {
  df[,column] <- as.numeric(df[,column])
}

str(df)

#remove NA
df <- df[-which(is.na(df$danceability)),]
```

```{r clustering}
scaled.data <- as.matrix(scale(df))
set.seed(123)
fviz_nbclust(scaled.data, kmeans, method="wss")

cluster <- kmeans(df, 2)

cluster.output <- data.frame(df,
                             cluster$cluster)

clusplot(df, cluster$cluster,
         color=T,
         shade=T,
         labels=0,
         lines=0)

write.csv(cluster.output, file="Cluster Combined Results.csv")
```

``` {r regression}
summary(lm(rank~., df))
```

```{r logit}
top10 <- c()

for (row in 1:length(rownames(df))) {
  if (df$rank[row] <= 10) {
    top10 <- c(top10, 1)
  } else {
    top10 <- c(top10, 0)
  }
}

table(top10)


df.logit <- data.frame(rank = music$rank,
                 year = music$year,
                 danceability = spotify$danceability,
                 energy = spotify$energy,
                 loudness = spotify$loudness,
                 speechiness = spotify$speechiness,
                 acousticness = spotify$acousticness,
                 instrmentalness = spotify$instrumentalness,
                 liveness = spotify$liveness,
                 tempo = spotify$temp,
                 duration = spotify$duration_ms,
                 girl = lyrics$girl,
                 love = lyrics$love, 
                 ooh = lyrics$ooh,
                 want = lyrics$want,
                 baby = lyrics$baby,
                 yeah = lyrics$yeah,
                 back = lyrics$back,
                 cause = lyrics$cause,
                 thing = lyrics$thing,
                 stringsAsFactors = F)

for (column in 1:length(colnames(df.logit))) {
  df.logit[,column] <- as.numeric(df.logit[,column])
}

str(df.logit)
#remove NA
df.logit <- df.logit[-which(is.na(df.logit$danceability)),]
df.logit <- data.frame(df.logit,
                       top10)
df.logit$top10 <- as.factor(df.logit$top10)

test <- df.logit[which(df.logit$year == 2015),
                 -which(colnames(df.logit) == "year")]
  actual <- df.logit[which(df.logit$year == 2015),
                 which(colnames(df.logit) == "top10")]
train <- df.logit[which(df.logit$year != 2015),
                 -which(colnames(df.logit) == "year")]

logit.model <- glm(top10~., train, family=binomial(link="logit"))
prediction <- predict(logit.model, test, type="response")

results <- c()
for (number in 1:length(prediction)) {
  if (prediction[number] > 0.5) {
    results <- c(results,1)
  } else {
    results <- c(results,0)
  }
}

table(results,actual)
```

```{r limit years}
df <- data.frame(rank = music$rank,
                 year = music$year,
                 danceability = spotify$danceability,
                 energy = spotify$energy,
                 loudness = spotify$loudness,
                 speechiness = spotify$speechiness,
                 acousticness = spotify$acousticness,
                 instrmentalness = spotify$instrumentalness,
                 liveness = spotify$liveness,
                 tempo = spotify$temp,
                 duration = spotify$duration_ms,
                 girl = lyrics$girl,
                 love = lyrics$love, 
                 ooh = lyrics$ooh,
                 want = lyrics$want,
                 baby = lyrics$baby,
                 yeah = lyrics$yeah,
                 back = lyrics$back,
                 cause = lyrics$cause,
                 thing = lyrics$thing,
                 stringsAsFactors = F)

for (column in 1:length(colnames(df))) {
  df[,column] <- as.numeric(df[,column])
}

str(df)

#remove NA
df <- df[-which(is.na(df$danceability)),]

limit5 <- df[which(df$year > 2010),
             -which(colnames(df) == "year")]
limit7 <- df[which(df$year > 2008),
             -which(colnames(df) == "year")]
limit10 <- df[which(df$year > 2004),
             -which(colnames(df) == "year")]
  
#regression
summary(lm(rank~., limit5))
summary(lm(rank~., limit7))
summary(lm(rank~., limit10))
```